---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# STA 326 2.0/ ASP 454 2.0 Programming and Data ANalysis with R

## Assignment 3 - 25 Marks


### Individual project: Coronavirus country report

We are still facing the global crisis of the COVID-19 pandemic. As individuals we can do our part in keeping others safe by following the rules suggested by the government and WHO. As part of the data science community we can also provide our contribution, by analysing the data and helping to communicate it. Such type of support is high demand among government and related institutions. The purpose of this assignment is to fulfil that role.


The project represents 25% of your final grade for STA 326 2.0/ ASP 454 2.0.

### 1. Guidelines

I have assigned a country for each of you. You need to write a report by analysing coronavirus data of the country that I have assigned. The report should be written in Rmarkdown and the output format should be HTML. All your reports will be published on the course website here: https://hellor.netlify.app

In order to view name of the country corresponds to your index number use the steps give below:

**Step 1**

Install `devtools` package.

**Step 2**

Type the following code to get access to the `sta3262` package. This package provides a function to view the country. Please note that you may have  `sta3262` already installed in to your machine. However, you should reinstall `sta3262` in order to get access to the assignment 3.

```r
library(devtools)
devtools::install_github("thiyangt/sta3262")

```

**Step 3**

Load the `sta3262` using the command

```{r, comment=NA}
library(sta3262)
```


**Step 4**

To view your country, 

```{r, comment=NA}
get_individual_project_country("AS2018000") # type your index n. within ""
```

**Please send me an email if your index number is not in the list or if the function outputs** character(0).

### 2. To obtain coronavirus data corresponds to your country


You can download coronavirus data from the coronavirus package in R. This help file here https://hellor.netlify.app/slides/coronavirus_assignment_help_file
gives some useful hints on how to extract and visualise relevant data from the package. 


### 3. Project content:

i. Brief introduction of the country:

-	Visualization of the location
-	Climate condition
-	Lock-down status of the country, etc.
-	Actions taken by the government
-	Etc.

ii. Exploratory Data Analysis

-	Visualization of death, confirmed, recovered, active cases
-	Comparison with other countries
-	etc

iii. Conclusions and Discussion

iv. References

Please click here to view the last year reports: https://hellor.netlify.app/coronareports/

You are not allowed to copy the content of this site.

### 4. Marks


Content: 20 Marks (Data Analysis and interpretation -15, Overall format - 5)

Participation in the LMS discussion forum: 5 Marks (You can participate by providing answers to your colleagues’ questions, providing links to useful reports/ blogposts, etc, and asking questions)

Please use the discussion forum on LMS to ask questions related to the project. 

### 5. Submission


Please submit .rmd file and .html file. Please name the files using your index number and the country.

Example: 2017000US.rmd and 2017000US.rmd



### 5. Deadline

28 September 2021, 3.00PM (IST)

### Other

Please be safe during this period and help one another. Your safety is my main priority. If you have any issues with submitting your assignment due to any unexpected circumstances, you can apply for an extension or special consideration. For special consideration or deadline extensions, please send me an email: ttalagala@sjp.ac.lk informing your request with the email title: STA 326 2.0 Individual project _deadline_extension_(your_Index_Number). Please refer to the course rules at https://hellor.netlify.app/2021/week0.html#10



