---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# STA 326 2.0/ ASP 454 2.0/ CSC 381 2.0 Programming and Data ANalysis with R

## Assignment 2 - 30 Marks


### Individual project: Coronavirus country report

The COVID-19 pandemic is not over yet. As individuals we can do our part in keeping others safe by following the rules suggested by the government and WHO. As part of the data science community we can also provide our contribution, by analysing the data and helping to communicate it. Such type of support is high demand among government and related institutions. The purpose of this assignment is to fulfil that role.


The project represents 30% of your final grade for STA 326 2.0/ ASP 454 2.0/ CSC 381 2.0.

### 1. Guidelines

I have assigned a country to each of you. You need to write a brief report by analyzing coronavirus data for the country that I have assigned. The report should be written in Rmarkdown and the output format should be HTML. All your reports will be published on the course website here: https://hellor.netlify.app

In order to view name of the country corresponds to your index number use the steps give below:

**Step 1**

Install `devtools` package.

```r
install.packages("devtools")
```

**Step 2**

Type the following code to get access to the `sta3262` package. This package provides a function to view the country. Please note that you may have  `sta3262` already installed in to your machine. However, you should reinstall `sta3262` in order to get access to the assignment 2.

```r
library(devtools)
devtools::install_github("thiyangt/sta3262")

```

**Step 3**

Load the `sta3262` using the command

```{r, comment=NA}
library(sta3262)
```


**Step 4**

To view your country, 

```{r, comment=NA, eval=FALSE}
get_individual_project_country("AS2018000") # type your index n. within ""
```

**Please send me an email if your index number is not in the list or if the function outputs** character(0).

### 2. To obtain coronavirus data corresponds to your country

```r
install.packages("coronavirus")
library(coronavirus)
```

You can download coronavirus data from the coronavirus package in R. This help file here https://hellor.netlify.app/slides/coronavirus_assignment_help_file
gives some useful hints on how to extract and visualise relevant data from the package. 


### 3. Project content:

i. Brief introduction of the country:

ii. Exploratory Data Analysis

-	Visualization of death, confirmed, recovered, active cases
-	Comparison with other countries
-	etc

iii. Conclusions and Discussion

iv. References

Please click here to view the last year reports: https://hellor.netlify.app/coronareports/

You are not allowed to copy the content of this site.

Please do not include any external images.

### 4. Marks


Content: 30 Marks (Data Analysis and interpretation 20, Overall format - 10)



Please use the discussion forum on LMS to ask questions related to the project. 

### 5. Submission


Please submit .rmd file and .html file. Please name the files using your index number and the country.

Example: 2017000US.rmd and 2017000US.rmd


### 5. Deadline

5 Oct 2022, 3.00PM (IST)


