---
title: "STA 326 2.0 Programming and Data Analysis with R"
subtitle: "Data Manipulation with dplyr"
author: "Dr Thiyanga Talagala"
date: "Online distance learning/teaching materials during the COVID-19 outbreak."
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(emo)
```


# packages


```{r, comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(magrittr)

```

---
# `dplyr` verbs

- `filter`

- `select`

- `mutate`

- `summarise`

- `arrange`

- `group_by`

---

# Dataset

```{r}
library(gapminder)
str(gapminder)
head(gapminder)
```

---

# `filter`

- Picks observations by their values.

- Takes logical expressions and returns the rows for which all are `TRUE`.

---

# `select`

- Pick variables by their names.

---

# `mutate`

- Create new variables with functions of existing variables

---

# `summarise`

- Collapse many values down to a single summary

---

# `arrange`

- Reorder the rows


---

# `group_by`
---
class: center, middle


Slides available at: hellor.netlify.app

All rights reserved by [Thiyanga S. Talagala](https://thiyanga.netlify.com/)



