---
title: "STA 326 2.0 Programming and Data Analysis with R ^[All rights reserved by Thiyanga S. Talagala]"
subtitle: "Exploring `iris` dataset with `qplot`"
date: "30 March 2020"
geometry: "left=3cm,right=3cm,top=2cm,bottom=2cm"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```


# Stage 1:  Planning your analysis

## Step 1: Dataset overview and description

Before we get started let's look at the data and plan the analysis.

**Load iris dataset**
```{r, comment=NA}
data(iris)
```

Here is a glimpse of the dataset

```{r, comment=NA}
head(iris)
```

We have four qunatitative variables: `r colnames(iris)[1:4]` and one qualitative variable `r colnames(iris)[5]`

## Step 2: Let's look at the graphs we could use to explore variables one by one.

Plots that could be used to to summarize quantitative variables

  - Box and whisker plot

  - Histograms

  - Dotplots 

  - Density plot
  
  - Stem and leaf displays
  
Note: Stem and leaf displays are best-suited for small to moderate datasets, whereas others such as histograms and Box and whisker plots are best-suited for large datasets. Graph types such as Box and whisker plots are good at depicting differences between distributions and identyfying outliers.

Plotss that could be used to to summarize qualitative variables

  - Bar chart

  - Pie chart

## Step 3: Next we will look at two variables at a time.

- Quantitative and Quantitative: Scatter plots

- Quantitative and Qualitative: 


   


