---
title: "Tidyverse - Script 1"
author: "Thiyanga Talagala"
date: "07/04/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Introduction to Tidyverse

## 1.1 What is tidyverse?

## 1.2 It is a collection of R packages

```{r, eval=FALSE}
library(tidyverse)
```

will load the core tidyverse packages in R. All packages seamlessly integrate and work together harmoniously. 

## 1.3 Tidyverse workflaw

## 1.4 Tidyverse workflaw with packages

In terms of R packages, the workflow is nicely depicted as in this picture,

**1.4.1 Importing  data**

**1.4.2 Tidyr**

*Organise rows of data into unique values.*

*Here, there's one row for each acoustic measurement so that each word is spread acros five rows.
This dataset is considered "tall" because there are many more rows than columns, giving the
impression of a tall rectangle when you view it.*

*Why are there different ways of storing the data? It depends on the research question you're asking.
If you're interested in each word, you might want to keep the extra-wide format so that each row
represents a word. If you're interested in each acoustic measurement individually, you might want
to use the tall version.*

*It's good to know how to convert your data from wide to tall and vice versa. Not only because it lets
you look at it in different ways, but because certain kinds of visualizations require one or the other.*

**1.4.3 Transform your data**

Once you have your data loaded, yo'll probably need to do some cleaning before moving on to analysis. We call this data pre-processing. 

This step is important to make sure that your dataset is nice and clean and consistently formatted and [ready for analysis.]. Doing this in base R is little bit harder.


*[Now that you've got your data loaded in], you'll proably need to do some preprocessing before moving
on to analysis. [This is also called cleaning your data.] Clean your data by removing columns, renaming columns, re-ordering columns, adding columns, etc.*

*In my experience, my R code always starts off with many lines of preprocessing to
make sure everything is nice and clean and consistently formatted before moving on to the statistics
and visualizations.*

*What do I mean by "processing"? Well, it's likely that your data isn't exactly the way it needs to be
for analysis. Maybe the column names aren't right, or a number that's supposed to be a number is
actually being treated like text, or you may want to add, remove, or reorder columns entirely. Even
if your data is pristine, you may have to do set a reference level for categorical data or filter the data
in some way. All of this can be done using functions in the dplyr package, one of the core
tidyverse packages. I'll cover some basics, but you can read more about the topics in this section
R for Data Science chapters 18 on Pipes (http://r4ds.had.co.nz/pipes.html), and Chapter 5 "Data
Transformation" (http://r4ds.had.co.nz/transform.html)*

**1.4.4: Visualising**

**1.4.5: Modelling**

**1.4.6: Communicate**

# Data

We will be using the gapminder dataset pioneered by Hans Rolling. These data represent the health and wealth of every nation in the world. 

# 2. Pipe function (%>%)

This little function is one of the most  useful feature feature of the tidyverse. 
What the pipe does is it takes the output of the function on the left and feeds it to the right function as its first argument. 


# 3. Tibble

# 4. Factor


----------------------------------------------------------------------------------------------------------------------------------
# Script 2: readr

We saw how to get your data into R using `read_csv` and `read_excel` and how these
functions are a bit more efficient than the base R equivalents.




