---
title: "STA 517 3.0 Programming and Statistical Computing with R"
subtitle: "Functionals"
author: "Dr Thiyanga Talagala"
date: ""
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
library(emo)
```

# Functionals

> A functional is a function that takes a function as an input and returns a vector as output.

> Hadley Wickham, Advanced R

```{r, comment=NA}
statistic <- function(f){
  data <- c(10, 20, 30, 40, 62, 63)
  f(data)
}

```

```{r, comment=NA}
statistic(mean)
statistic(sum)
```


---

# Use of functionals: lapply

lapply: loop over a list and evaluate a function on each element.

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA}
#We are passing `mean` is an argument to lapply
lapply(x, mean)
lapply(x, sum)
```


---
# Use of functionals: lapply (cont.)

```{r, comment=NA, message=FALSE}
cv <- function(data){sd(data)/mean(data)}
lapply(x, cv)
```

---
# lapply is a for-loop replacement

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA, message=FALSE}
result_x <- list()
result_x
for (i in 1:2){
 result_x[[i]] <- mean(x[[i]])
  
}
result_x
```

---
# Use of functionals: sapply

lapply: loop over a list and evaluate a function on each element.

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

```

```{r, comment=NA}
#We are passing `mean` is an argument to lapply
sapply(x, mean)
sapply(x, sum)
```

Same as `lapply` but the output is a vector.
---

# `map()` function in purrr

```{r, comment=NA}
library(purrr)
```

```{r, comment=NA}
x <- list( a = 1:8, b = c(2.1, 3.2, 4.2, 5, 6))
x

map(x, mean)

```

- The base equivalent to `map()` is `lapply()`.

---
# `map` syntax

> map(.x, .f)

`.x` - The object we want to iterate over (a vector, list or dataframe)

`.f` - function (What are we going to do?)

For each element (vector/ list) or for each column in a data frame apply `.f` function.

```{r, comment=NA}
map(c(4, 9, 16), sqrt)
```
---

## Demo: Visualization of `map`


---

# `map()` with data frames

```{r}
head(trees)
trees %>% 
  map(mean)
```



---
# `map()` with data frames

```{r}
iris %>% 
  dplyr::select_if(is.numeric) %>% 
  map(mean)
```


---

# `map` additional inputs to the function

Eg: `na.rm=TRUE`

### Method 1

```{r}
abc <- list(a = c(1, NA, 3), b = 4:6, c = 10:12)
map(abc, mean)
map(abc, mean, na.rm=TRUE)
```

---

### Method 2

```{r}
map(abc, function(.x){
  mean(.x, na.rm=TRUE)
})

```


---
### Method 3

```{r}
map(abc, ~mean(.x, na.rm=TRUE))

```


---
## Your turn

**Question 1**

Identify the number of unique values in each column of the `gapminder` dataset

```{r, echo=FALSE}
library(countdown)
countdown(minutes = 5L, seconds = 00)
```


---
## Returning types 

- `map`: list

- `map_chr` : charactor vector

- `map_dbl` : double vector

- `map_int` : integer vector

- `map_lgl` : logical vector

- `map_dfc` : data frame (by column)

- `map_dfr` : data frame (by row)

---

```{r, echo=FALSE, comment=NA, message=F, warning=FALSE}
library(tidyverse)
```

```{r}
abc <- list(a = c(1, NA, 3), b = 4:6, c = 10:12)
abc %>% map(is.numeric)
abc %>% map_lgl(is.numeric)
abc %>% map_chr(is.numeric)
```

---

```{r, comment=NA}
map_output <- map(mtcars, function(x) length(unique(x)))
head(map_output, 3)
```

---

```{r, comment=NA}
set.seed(2020)
x <- list(a=rnorm(5), b=rnorm(6))
map(x, mean)
```


```{r, comment=NA}
map_df(x, mean)
```

---

**Question 2**

Identify the number of unique values in each column of the `gapminder` dataset. The output of the map function should be a integer vector.

```{r, echo=FALSE}
library(countdown)
countdown(minutes = 5L, seconds = 00)
```


---
## Your turn

**Question 3**

Split the iris data set according to species type and fit simple linear regression model between sepal.length and sepal.width



```{r, echo=FALSE}
library(countdown)
countdown(minutes = 5L, seconds = 00)
```
---
# map2(.x, .y, .f)

For each element of `.x` and `.y` do `f(.x, .y)`

```{r}
abc
cde <- list(x = c(10, 20, 30), y= c(100, 200, 300), z=c(0, 0, 0))
cde
```



---
# map2(.x, .y, .f)

```{r}
map2(abc, cde, sum)
```


---

## Your turn

**Question 4**

Split the iris data set according to species type and fit simple linear regression model between sepal.length and sepal.width **and obtain the predictions.**



```{r, echo=FALSE}
library(countdown)
countdown(minutes = 5L, seconds = 00)
```


---

# `pmap`

take more than 2 lists or data frame with argument names

```{r}
pqr <- list(p=c(1, 1, 1), q=c(2, 2, 2), r =c(3, 3, 3))
l <- list(abc, cde, pqr)
pmap(l, sum)
```

If you want to operate over more than 3 inputs you always need to first include the elements into a list




---

class: center, middle



Slides available at: hellor.netlify.app

All rights reserved by [Thiyanga S. Talagala](https://thiyanga.netlify.com/)


Reference: Advanced R, Hadley Wickham
