---
title: "STA 326 2.0 Programming and Data Analysis with R"
subtitle: "Lesson 3: Functions in R"
author: "Dr Thiyanga Talagala"
date: "2020-02-24"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: 
      - default
      - default-fonts
      - duke-blue
      - hygge-duke
      - libs/cc-fonts.css
      - libs/figure-captions.css
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpans: true
      countIncrementalSlides: false
---
```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

## Functions in R

ðŸ‘‰ðŸ» Perform a specific task according to a set of instructions.

--
ðŸ‘‰ðŸ»  Some functions we have discussed so far,

> `c`, `matrix`, `array`, `list`, `data.frame`, `str`, `ndim`, `length`, `nrow`, `plot`

--

ðŸ‘‰ In R, functions are **objects** of **class** *function*.

```{r, comment=NA}
class(length)
```

--

ðŸ‘‰ðŸ» There are basically two types of functions:

    
 > ðŸ’» Built in functions 
  
      Already created or defined in the programming framework to make our work easier.
> ðŸ‘¨ User defined functions
           
      Sometimes we need to create our own functions for a specific purpose.

---

## Basic components of a function

.pull-left[

## Syntax

```r

name <- function(arg1, aug2, ...){

<FUNCTION BODY>

return(value)

}

```
]

.pull-right[

## Example

```{r}
calSqrt <- function(x){

a <- x^2
b <- x^3
out <- c(a, b)
names(out) <- c("squared", "cubed")
out # or return(out)

}
```

## Evaluation

```{r, comment=NA}
calSqrt(2)
```

]



---


---
class: center, middle

Slides available at: hellor.netlify.com

[CC BY-NC license](https://creativecommons.org/licenses/by-nc/4.0/)


