---
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, message = FALSE, warning = FALSE)
```

## R packages

```{r}
library(tidyverse)
library(skimr)
library(knitr)
library(naniar)
library(ggrepel)
library(recipes)
library(scales)
theme_set(theme_bw())
#theme_set(theme_light())
#library(extrafont) #Import extra font
#extrafont::font_import()
```

## About data

This is a dataset of individual people who were freed from slave ships during the last few decades. This gives information about enslaved people in America.

## Read and Skim data

```{r}
african_names <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-06-16/african_names.csv")
glimpse(african_names)
skimr::skim(african_names)

```

# Exploratory data analysis

## Where these ships were disembarked?

```{r}
african_names %>%
  count(port_disembark, sort = TRUE) %>%
  kable()
```

## Where these ships were embarked?

```{r}
african_names %>%
  count(port_embark, sort = TRUE) %>%
  kable()
```



## Note: `add_count`

```{r}
#count the unique values of one or more variables
iris %>% add_count(Species) %>% head(3)
```

## Port of embark by year

```{r}

african_names %>%
  add_count(port_embark) %>%
  mutate(port_embark = case_when(
    n < 4000 ~ "Other",
    TRUE ~ port_embark
  )) %>%
  ggplot(aes(port_embark, year_arrival, fill = port_embark)) +
  geom_boxplot(alpha = 0.4, show.legend = FALSE) +
  labs(x = NULL, y = "Year")
```

## Distribution of years of arrivals

```{r}
african_names %>%
  ggplot(aes(year_arrival)) +
  geom_histogram(bins = 20, fill = "midnightblue", alpha = 0.7, col="white") +
  scale_y_continuous(labels = scales::comma_format()) +
  labs(
    y = "Number of liberated individuals",
    x = "Year"
  )
```

## Handling missing values

```{r}
african_names %>% gg_miss_var()
```



```{r}
african_names %>%
  select(gender, age, height, year_arrival) %>%
  gg_miss_upset()
```

## Relationship between years of arrival and gender

```{r}
african_names %>%
  ggplot(aes(gender, year_arrival, fill = gender)) +
  geom_boxplot(alpha = 0.4, show.legend = FALSE) +
  labs(x = NULL, y = "Year")
```

## Relationship between age and gender

```{r}
african_names %>%
  ggplot(aes(gender, age, fill = gender)) +
  geom_boxplot(alpha = 0.4, show.legend = FALSE) +
  labs(x = NULL, y = "Age")
```

## Relationship between age and years of arrivals

```{r}
african_names %>%
  filter(year_arrival < 1850) %>%
  group_by(year_arrival) %>%
  summarise(age = mean(age, na.rm = TRUE)) %>%
  ggplot(aes(year_arrival, age)) +
  geom_line(alpha = 0.6, size = 1.5) +
  geom_smooth(method = "lm") +
  scale_y_continuous(limits = c(0, NA)) +
  labs(x = NULL, y = "Mean age")

```

## Relationship between age, year of arrivals and names of transported captives

```{r}
african_names %>%
  group_by(name) %>%
  summarise(
    n = n(), # how many people have that name
    age = mean(age, na.rm = TRUE),
    year_arrival = mean(year_arrival, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(-n) %>%
  filter(n > 30) %>%
  ggplot(aes(year_arrival, age)) +
  geom_text_repel(aes(label = name), size = 3) +
  geom_point(aes(size = n), color = "midnightblue", alpha = 0.7) +
  labs(
    x = "Mean year of arrival", y = "Mean age",
    size = "Number of people",
    title = "Age and year of arrival for most common names of transported captives",
    caption = "African Names Database from slavevoyages.org"
  )

```

## Impute missing values

```{r}
liberated_df <- african_names %>%
  filter(year_arrival < 1850) %>% 
  mutate(gender = case_when(
    gender == "Boy" ~ "Man",
    gender == "Girl" ~ "Woman",
    TRUE ~ gender
  )) %>%
  mutate_if(is.character, factor)

liberated_df

```

## Missing value imputation with `recipes`

```{r}
impute_rec <- recipe(year_arrival ~ gender + age + height, data = liberated_df) %>%
  step_meanimpute(height) %>%
  step_modeimpute(gender) %>%
  step_medianimpute(age)
imputed <- prep(impute_rec) %>% juice()
imputed
```

## Linear regression model

```{r}
fit_lm <- lm(year_arrival ~ gender + age, data = imputed)
summary(fit_lm)
tidy(fit_lm) %>%
  kable(digits = 3)
```

Acknowledgement: https://juliasilge.com/blog/captive-africans-voyages/
