---
title: "Untitled"
author: "Hasarali"
date: "2022-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r, warning=FALSE}
library(devtools)
devtools::install_github("thiyangt/sta3262")
library(sta3262)
get_individual_project_country("AS2019367")
```
# Introduction
    ##Belgium, country of northwestern Europe. It is one of the smallest and most densely populated European countries. Culturally, Belgium is a heterogeneous country straddling the border between the Romance and Germanic language families of western Europe.The country has a total of 860 miles (1,385 km) of land boundaries with neighbours; it is bounded by the Netherlands to the north, Germany to the east, Luxembourg to the southeast, and France to the south. Belgium also has some 40 miles (60 km) of shoreline on the North Sea


    
 
# Covid-19 pandemic in Belgium
   ## Soon after SARS-CoV-2 emerged in late 2019, Belgium was confronted with a first COVID-19 wave in March-April 2020. SARS-CoV-2 circulation declined in the summer months (late May to early July 2020). Following a successfully trumped late July-August peak, COVID-19 incidence fell slightly, to then enter two successive phases of rapid incline: in the first half of September, and then again in October 2020. The first of these coincided with the peak period of returning summer travelers; the second one coincided with the start of higher education’s academic year. The largest observed COVID-19 incidence occurred in the period 16–31 October, particularly in the Walloon Region, the southern, French-speaking part of Belgium. We examine the potential association of the higher education population with spatio-temporal spread of COVID-19, using Bayesian spatial Poisson models for confirmed test cases, accounting for socio-demographic heterogeneity in the population. We find a significant association between the number of COVID-19 cases in the age groups 18–29 years and 30–39 years and the size of the higher education student population at the municipality level. These results can be useful towards COVID-19 mitigation strategies, particularly in areas where virus transmission from higher education students into the broader community could exacerbate morbidity and mortality of COVID-19 among populations with prevalent underlying conditions associated with more severe outcomes following infection.
 
 
  
```{r, warning=FALSE}
library(tidyverse)
library(magrittr)
library(ggplot2)
library(plotly)
library(tidyverse)
library(coronavirus)
```
```{r, warning=FALSE}
data("coronavirus")
belgium_corona <- coronavirus %>% filter(country == "Belgium")
```
# Location of Belgium in World map
```{r, warning=FALSE}
library(maptools)
library(tibble)
library(ggrepel)
library(png)
library(grid)
library(sp)
data(wrld_simpl)

p <- ggplot() +
  geom_polygon(
    data = wrld_simpl,
    aes(x = long, y = lat, group = group), fill = "light blue", colour = "white"
  ) +
  coord_cartesian(xlim = c(-180, 180), ylim = c(-90, 90)) +
  scale_x_continuous(breaks = seq(-180, 180, 120)) +
  scale_y_continuous(breaks = seq(-90, 90, 100))

p +
  geom_point(
    data = belgium_corona, aes(x = long, y = lat), color = "red", size
    = 1
  )
```
# Explotary Data Analysis 
 
```{r, warning=FALSE}
 belgium_corona <- as.data.frame(belgium_corona)
head(belgium_corona)

```

```{r, warning=FALSE}
tail(belgium_corona)
```
```{r,, warning=FALSE}
summary(belgium_corona)
```
The table belgium_corona is consist of the no.of confirmed, death and recovered corona patients from 2020-01-22 to 2022-06-23 of the country belgium and the location of the country in the wordls map with its latitude and longitude.

```{r, warning=FALSE}
confirmed_cases <- filter(belgium_corona, type == "confirmed")

ggplot(confirmed_cases,aes(x=date,y=cases)) + geom_path(size=1) + labs(title = "Confirmed cases in Belgium")
```
Above graph showas the confirmed no.of corona patients in belgium in daily basis. From 2021 to 2022, the number of confirmed cases will increase slightly and again in the middle of 2022, the number of confirmed cases will decrease. This shows the lack of control of the situation.
```{r, warning=FALSE}
death_cases <- filter(belgium_corona, type == "death")

ggplot(death_cases,aes(date, cases)) + geom_path(size=1) + labs(title = "Death cases in Belgium")
```

Above graph shows the no.of deadth from Corona in colombia in daily basis. This graph also shows that no.of death is decreasing from end of the 2020to 2022. Although fluctuations shows that in some days no.of deaths are greater than the previous day, still the no.of deaths per day has decresed than the previous week.
```{r, warning=FALSE}
recovered_cases <- filter(belgium_corona, type == "recovered")

ggplot(recovered_cases, aes(date, cases)) + geom_path(size=1) + labs(title = "Recovered cases in Belgium")
```

Above graph shows the no.of patients recoverd from corona in belgium in daily basis. No.of patients recovered in a day has zero.

```{r, warning=FALSE}
ggplot(belgium_corona, aes(date, cases)) + geom_path(size=1) + facet_grid(rows = vars(type)) +
  labs(title = "Confirmed, Death and Recovered cases in Belgium")

```


```{r, warning=FALSE}
ggplot(belgium_corona, aes(x = date, y = cases, colour = type)) + geom_path(size=1) + 
  labs(title="Confirmed, Death and Recovered cases in Belgium")

```
```{r, warning=FALSE}
belgium_corona <-  belgium_corona%>%
  group_by(type)%>%
  mutate(cum_cases = cumsum(cases))
```
```{r, warning=FALSE}
unique(coronavirus$country)
netherlands_corona<-coronavirus%>%filter(country=="Netherlands")

germany_corona<-coronavirus%>%filter(country=="Germany")

belgium_corona<-coronavirus%>%filter(country=="Belgium")
france_corona<-coronavirus%>%filter(country=="France")


countries_corona<-as.data.frame(rbind(netherlands_corona,germany_corona,belgium_corona,france_corona))
```
Above graph shows the no.of confirmed, death and recovered corona patients in belgium in daily.
```{r,, warning=FALSE}
countries_death<-countries_corona%>%filter(type=="death")
ggplot(countries_death,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+geom_point()+ 
  facet_wrap(~country,ncol = 4)+coord_flip()+
  ggtitle("Comparison  of covid-19 deaths between the neighbooring countries")
```
```{r, warning=FALSE}
countries_cases<-countries_corona%>%filter(type=="confirmed")

ggplot(countries_cases,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+geom_point()+ 
  facet_wrap(~country,ncol = 4)+coord_flip()+
  ggtitle("Comparison between the confirmed covid-19 
          patients between the neighbooring countries")
```
```{r, warning=FALSE}
countries_cases<-countries_corona%>%filter(type=="confirmed")
```
```{r, warning=FALSE}
ggplot(countries_cases,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+geom_point()+ 
  facet_wrap(~country,ncol = 4)+coord_flip()+
  ggtitle("Comparison between the confirmed covid-19 
          patients between the neighbooring countries")
```

# Conclusions and Discussion

       Compared to other neighbouring countries, the number of confirmed corona patients in Belgium is low and the recovery rate of corona patients is also low compared to other countries, but the death rate of corona patients is higher than the Netherlands.In the belgium corona virus data set there was a negative value of confirmed cases,but it is not significant so i have do conduct a data preprocessing step.

# References
    https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0264516#:~:text=During%20the%20Spring%202020%20COVID,with%20less%20fatalities%20%5B28%5D  
    https://en.m.wikipedia.org/wiki/COVID-19_pandemic_in_Belgium 
    https://github.com/thiyangt/CoronaSriLanka/blob/master/coronaSLDashboard.Rmd


 






