---
title: "UK Corona Analysis"
author: "AS2019315"
date: "2022-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

United Kingdom is an island country spanning an archipelago including Great Britain, located in Western Europe comprising England, Scotland, Wales, and Northern Ireland. United Kingdom is surrounded by the Atlantic Ocean, the North Sea, the English Channel, and the Irish Sea
```{r message=FALSE, warning=FALSE, include=FALSE}

library(tidyverse) 
library(magrittr)
library(coronavirus)
uk_corona <- coronavirus %>% filter(country == "United Kingdom")
sl_corona <- coronavirus %>% filter(country == "Sri Lanka")
chin_corona <- coronavirus %>% filter(country == "China")
us_corona <- coronavirus %>% filter(country == "US")
library(ggplot2) 
library(maptools) 
library(tibble) 
library(ggrepel) 
library(png) 
library(gridBase) 
library(sp) 
data(wrld_simpl)




```
# Location of uk

The United Kingdom of Great Britain and Northern Ireland (UK) is an island country that sits north-west of mainland Europe. It is made up of mainland Great Britain (England, Wales and Scotland) and the northern part of the island of Ireland (Northern Ireland). It has numerous smaller islands

```{r echo=FALSE, message=FALSE, warning=FALSE}
uk_location <- tibble(lat=55.3781   , long=3.436) 
p <- ggplot() + geom_polygon(data = wrld_simpl, aes(x = long, y = lat, group = group), fill = "gray", colour = "white") +
  
coord_cartesian(xlim = c(-180, 180), ylim = c(-90, 90)) +   
  scale_x_continuous(breaks = seq(-180, 180, 120)) +   
  scale_y_continuous(breaks = seq(-90, 90, 100)) 
 
p +   
  geom_point(data = uk_location, aes(x = long, y = lat), color = "red", size = 1)

```








# EXPLONATARY DATA ANALYSIS

 
```{r message=FALSE, warning=FALSE, include=FALSE}



eng_corona_conf <- uk_corona %>% filter(type=="confirmed")


eng_corona_rec <- uk_corona %>% filter(type=="recovery")


eng_corona_death <- uk_corona %>% filter(type=="death")


uk_corona_new <- uk_corona %>% mutate(cases=replace(cases,which(cases<0)
                                                    ,NA))
sl_corona_new <- sl_corona %>% mutate(cases=replace(cases,which(cases<0)
                                                    ,NA))
us_corona_new <- us_corona %>% mutate(cases=replace(cases,which(cases<0)
                                                    ,NA))
chin_corona_new <- chin_corona %>% mutate(cases=replace(cases,which(cases<0)
                                                    ,NA))








```

death,recovery, confirmed

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(plotly)

plot_ly(data= uk_corona_new,
        x=~date, 
        y=~cases, 
        type = 'scatter',
        mode = 'lines',
        color = ~type)
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(eng_corona_conf, aes(x=date, y=cases)) + geom_line() + ggtitle("UK: Daily Covid-19 Confirmed cases")


```



```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(eng_corona_death, aes(x=date, y=cases)) + geom_line() + ggtitle("UK: Daily Covid-19 Death cases")

```

## Comparison with Other Countries

```{r message=FALSE, warning=FALSE, include=FALSE}
e1 <- uk_corona_new %>% select(date,cases,type,country)
s1 <- sl_corona_new %>% select(date,cases,type,country)
u1 <- us_corona_new %>% select(date,cases,type,country)
c1 <- chin_corona_new %>% select(date,cases,type,country)

all <- rbind(e1,s1,u1,c1)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
plot_ly(data= all,
        x=~date, 
        y=~cases, 
        type = 'scatter',
        mode = 'lines',
        color = ~country)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
ggplot(all, aes(x=country, y=cases, col= type)) + geom_boxplot(outlier.shape = NA, alpha = 0.2) + geom_jitter(aes(col=type)) + ggtitle("A boxplot of type of cases with Other countries")

```

# DISCUSSION

March 6, 2020, the first death as a result of coronavirus was recorded in the UK.
the number of deaths in the UK has increased significantly since then.
On April 21, 1172 deaths were recorded, which is the highest total in single day in the UK since the outbreak began.

The percentage of patients who have recovered from the coronavirus in the UK currently stands at 2 per cent.





# REFERENCES
https://globaledge.msu.edu/countries/united-kingdom#:~:text=United%20Kingdom%20is%20an%20island,Channel%2C%20and%20the%20Irish%20Sea.

https://thecommonwealth.org/our-member-countries/united-kingdom#:~:text=The%20United%20Kingdom%20of%20Great,It%20has%20numerous%20smaller%20islands.

