---
title: "Filter observations"
output: 
  learnr::tutorial:
    progressive: true
    allow_skip: true
runtime: shiny_prerendered
description: >
  Learn how to filter observations in a data frame. Use `filter()` to extract
  observations from a data frame, and use `&`, `|`, and `!` to write logical
  tests.
---

```{r setup, include=FALSE}
library(learnr)
tutorial_options(exercise.timelimit = 60)
knitr::opts_chunk$set(error = TRUE)
```

## Welcome

This is a demo tutorial. Compare it to the [source code](https://github.com/rstudio/learnr/tree/master/inst/tutorials/ex-data-filter/ex-data-filter.Rmd) that made it.

###  

In this tutorial, you will learn how to:

* use `filter()` to extract observations from a data frame or tibble
* write logical tests in R
* combine logical tests with Boolean operators
* handle missing values within logical tests

The readings in this tutorial follow [_R for Data Science_](http://r4ds.had.co.nz/), section 5.2. 

### Prerequisites

To practice these skills, we will use the `flights` data set from the nycflights13 package. This data frame comes from the US [Bureau of Transportation Statistics](http://www.transtats.bts.gov/DatabaseInfo.asp?DB_ID=120&Link=0) and contains all `r format(nrow(nycflights13::flights), big.mark = ",")` flights that departed from New York City in 2013. It is documented in `?flights`.

We will also use the ggplot2 package to visualize the data. 

If you are ready to begin, click on!

## Filter rows with `filter()`

### filter()

`filter()` lets you use a logical test to extract specific rows from a data frame. To use `filter()`, pass it the data frame followed by one or more logical tests. `filter()` will return every row that passes each logical test.

So for example, we can use `filter()` to select every flight in flights that departed on January 1st. Click Run Code to give it a try:

```{r}
1+2
```