---
title: "Covid-19 in Morocco"
author: "IGDMW KUMARA"
date: "2022-10-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, comment=NA, echo=FALSE, warning=FALSE, force=TRUE,eval=FALSE}
#install.packages("devtools")
library(devtools)
devtools::install_github("thiyangt/sta3262")
library(sta3262)
get_individual_project_country("AS2019423")
#install.packages("coronavirus")
library(coronavirus)

```
# **Introduction**

Morocco is a country located in North Africa. It has a coast on the Atlantic Ocean that reaches past the Strait of Gibraltar into the Mediterranean Sea. It is bordered by Spain, Algeria, and Western Sahara. Much of Morocco is mountainous. The Atlas Mountain and Rif Mountains are located on the northern coast and interior. The government system is a parliamentary constitutional monarchy; the chief of state is the king, and the head of government is the prime minister. Morocco is a member of the League of Arab States (Arab League).

## _Location of Morocco_

```{r, message=FALSE, echo=FALSE}
library(coronavirus)
library(tidyverse)
library(magrittr)
coronavirus$cases <- abs(coronavirus$cases)
coronavirus <- as.data.frame(coronavirus, stringAsFactors = TRUE)
morocco_corona <- coronavirus %>% filter(country == "Morocco")

library(ggplot2)
library(maptools)
library(tibble)
library(tidyverse)
library(ggrepel)
library(png)
library(grid)
library(sp)
data("wrld_simpl")


p <- ggplot() +
  geom_polygon(
    data = wrld_simpl,
    aes(x = long, y = lat, group = group), fill = "gray", colour = "white"
  ) +
  coord_cartesian(xlim = c(-180, 180), ylim = c(-90, 90)) +
  scale_x_continuous(breaks = seq(-180, 180, 120)) +
  scale_y_continuous(breaks = seq(-90, 90, 100))

p +
  geom_point(
    data = morocco_corona , aes(x = long, y = lat), color = "red", size
    = 1
  )
```



## _Situation of Covid-19 in Morocco_

The COVID-19 pandemic in Morocco is part of the worldwide pandemic of coronavirus disease 2019 (COVID-19) caused by severe acute respiratory syndrome coronavirus 2 (SARS-CoV-2). The virus was confirmed to have spread to Morocco on 2 March 2020, when the first case COVID-19 case was confirmed in Casablanca.[3] It involved a Moroccan expatriate residing in Bergamo, Italy, who arrived from Italy on 27 February. A second case was confirmed later that same day involving an 89-year-old woman Moroccan residing in Italy who had returned to Morocco on 25 February from Bologna, Italy.[4] As the outbreak widened in Morocco, in mid-March the Government closed schools and suspended international passenger flights.

## _Impacts of Covid19 in Morocco_

The COVID-19 pandemic has hit the Moroccan economy hard, as elsewhere in the world. A collapse in external demand and a lockdown lasting more than three months have profoundly altered economic activity in Morocco, causing its first recession since 1995. The implementation of the confinement and social distancing measures was strict and came two weeks after the detection of the first cases of COVID-19 in Morocco on March 2, 2020. The lockdown was extended three times and lasted around 14 weeks. This strategy prevented a sharp increase in cases in the early part of the pandemic, and enabled the preparation of health infrastructure to deal with potentially more severe waves of contamination. However, the price was the consequent economic downturn.

# **Exploratory Data Analysis**

## Visualization of death, confirmed, recovered, active cases

```{r, comment=NA, echo=FALSE, warning=FALSE, eval=FALSE}
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)
 coronavirus
 summary(coronavirus) #summery of whole data set of Morocco
 
  confirmed_cases_morocco_corona <- morocco_corona %>% filter(type=="confirmed") 
  sum(confirmed_cases_morocco_corona$cases) #Total number of confirmed cases
  
  deaths_morocco_corona <- morocco_corona_new %>% filter(type=="death") 
  sum(deaths_morocco_corona$cases) #Total number of deaths
  
  recovered_cases_morocco_corona <- morocco_corona %>% filter(type=="recovery")
  sum(recovered_cases_morocco_corona$cases) #Total number of recovered cases
  
  
```

_Confirmed cases in Morocco_

```{r, comment=NA, echo=FALSE, warning=FALSE}
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)
 
 confirmed_cases_morocco_corona <- morocco_corona %>% filter(type=="confirmed")
 summary(confirmed_cases_morocco_corona) #summary of confirmed cases in Morocco
 
 ggplot(confirmed_cases_morocco_corona, aes(x=date, y=cases)) +geom_point() + ggtitle("Covid-19 Confirmed cases in Morocco") #plot of confirmed cases in Morocco
 
```

_Deaths in Morocco_

```{r, comment=NA, echo=FALSE, warning=FALSE}
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)
 
 deaths_morocco_corona <- morocco_corona %>% filter(type=="death")
 summary(deaths_morocco_corona) #summary of death cases in Morocco
 
 ggplot(deaths_morocco_corona, aes(x=date, y=cases,)) +geom_point() + ggtitle("Covid-19 Deaths in Morocco") #plot of deaths in Morocco

```

_Recovered cases in Morocco_

```{r, comment=NA, echo=FALSE, warning=FALSE}
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)
 
 recovered_cases_morocco_corona <- morocco_corona %>% filter(type=="recovery")
 summary(recovered_cases_morocco_corona) #summary of recovered cases in Morocco
 
 
 recovered_cases_morocco_corona <- recovered_cases_morocco_corona %>% mutate(cases=replace(cases,which(cases < 0),NA))
 
 
 ggplot(recovered_cases_morocco_corona, aes(x=date, y=cases)) +geom_point() + ggtitle("Covid-19 Recovered cases in Morocco")

```

_Active cases in Morocco_

```{r, comment=NA, echo=FALSE, warning=FALSE}

library(coronavirus)
morocco_corona <- coronavirus %>% filter(country == "Morocco")
morocco_corona2<- morocco_corona %>% filter(cases >= 0)
corona_morocco_wider <- morocco_corona2 %>% pivot_wider(names_from=type, values_from=cases)
corona_morocco_wider$active_cases_morocco_corona <- corona_morocco_wider$confirmed - corona_morocco_wider$death - corona_morocco_wider$recovery #total number of the active cases in Morocco


ggplot(corona_morocco_wider, aes(x=date, y=active_cases_morocco_corona)) + geom_point() + ggtitle("Covid-19 Active cases in Morocco")

 
```

_Covid-19 cases in Morocco by type_

```{r,comment=NA, echo=FALSE, warning=FALSE}

library(ggplot2)

ggplot(morocco_corona, aes(x=type, y=cases, fill=type)) + geom_col() + ggtitle("Covid-19 Confirmed cases by type") + ylim(0,2500)

```

## Comparison with other countries

```{r, comment=NA, echo=FALSE, warning=FALSE}

 unique(coronavirus$country)
 
```

_Comparison of deaths between Morocco, Spain and Algeria_

```{r, comment=NA, echo=FALSE, warning=FALSE}
 
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)

 morocco_corona<-coronavirus%>%filter(country=="Morocco")
 
 spain_corona<-coronavirus%>%filter(country=="Spain")
 
 algeria_corona<-coronavirus%>%filter(country=="Algeria")
 

 neighbor_corona<-as.data.frame(rbind(morocco_corona,algeria_corona,spain_corona))
 summary(neighbor_corona) #summary of all cases 2 in neighboring countries
 
 
 neighbor_corona <- neighbor_corona %>% mutate(cases=replace(cases,which(cases < 0),NA))
 
 
 deaths_neighbor_corona <- neighbor_corona %>% filter(type=="death")
summary(deaths_neighbor_corona)

ggplot(deaths_neighbor_corona,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+ 
  facet_wrap(~country,ncol = 5)+coord_flip()+
  ggtitle("Comparison  of death cases between botswana & 2 neighbor countries")

```


_Comparison of confirmed cases between Morocco, Spain and Algeria_

```{r, comment=NA, echo=FALSE, warning=FALSE}
 
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)

 morocco_corona<-coronavirus%>%filter(country=="Morocco")
 
 spain_corona<-coronavirus%>%filter(country=="Spain")
 
 algeria_corona<-coronavirus%>%filter(country=="Algeria")
 

 neighbor_corona<-as.data.frame(rbind(morocco_corona,algeria_corona,spain_corona))
 summary(neighbor_corona) #summary of all cases 2 in neighboring countries
 neighbor_corona <- neighbor_corona %>% mutate(cases=replace(cases,which(cases < 0),NA))
 
 confirmed_cases_neighbor_corona <- neighbor_corona %>% filter(type=="confirmed")
 summary(confirmed_cases_neighbor_corona)

ggplot(confirmed_cases_neighbor_corona,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+ 
  facet_wrap(~country,ncol = 5)+coord_flip()+
  ggtitle("Comparison  of confirmed cases between botswana & 2 neighbor countries")

```


_Comparison of recovered cases between Morocco, Spain and Algeria_

```{r, comment=NA, echo=FALSE, warning=FALSE}
 
library(ggplot2)
library(coronavirus)
library(magrittr)
library(tidyverse)
library(dplyr)
library(tidyr)
library(tibble)
library(knitr)

 morocco_corona<-coronavirus%>%filter(country=="Morocco")
 
 spain_corona<-coronavirus%>%filter(country=="Spain")
 
 algeria_corona<-coronavirus%>%filter(country=="Algeria")
 

 neighbor_corona<-as.data.frame(rbind(morocco_corona,algeria_corona,spain_corona))
 summary(neighbor_corona) #summary of all cases 2 in neighboring countries
 
  neighbor_corona <- neighbor_corona %>% mutate(cases=replace(cases,which(cases < 0),NA))
 
 recovered_cases_neighbor_corona <- neighbor_corona %>% filter(type=="recovery")
summary(recovered_cases_neighbor_corona)

ggplot(recovered_cases_neighbor_corona,
       aes(x=date,y=cases,fill=type,col=country))+
  geom_line()+ 
  facet_wrap(~country,ncol = 5)+coord_flip()+
  ggtitle("Comparison  of death cases between botswana & 2 neighbor countries")

```

There are some missing values in here.

# **Conclusions and Discussion**
This data set was loaded by corona virus package in R Markdown.There were 1,195,883 confirmed cases were reported in Morocco and 16,094 cases out of it were deaths and 586,579 patients were recovered. And there are 593,210 cases are still active.There were no any negative values for confirmed cases and death cases of Morocco, but there was negative values for recovered cases and it was corrected by mean imputation interpolation method.The data were not categorized into provinces, therefore we can't analyze in province wise.


# **References**

https://globaledge.msu.edu/countries/morocco
https://en.wikipedia.org/wiki/COVID-19_pandemic_in_Morocco
https://www.policycenter.ma/publications/impact-covid-19-morocco-macroeconomic-sectoral-and-regional-effects