---
title: "STA 326 2.0 Programming and Data Analysis with R"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Tutorial 1 - Answers

### Create objects

1. 

2. Generate a sequence using the code `seq(from=1, to=10, by=1)`. What other ways can you generate the same sequence?


3. Using the function  `rep()` , create the below sequence `1, 2, 3, 4, 1, 2, 3, 4, 1, 2, 3, 4`


### Subsetting

4. Consider the vector 

```{r, comment=NA}
x <- c(80, 39, NA, 51, 51, 11, NA, NA, NA, 100, 80)
```

Write an R code to extract non-missing values in `x`

```{r, comment=NA}
# Answer 4
x[!is.na(x)]

```

Write an R code to extract missing values and odd-numbers in `x`

```{r, comment=NA}
x[x %% 2 == 1]
```

Write an R code to extract odd numbers on `x`

```{r, comment=NA}
y <- x[x %% 2 == 1]
y[!is.na(y)]
```

5. Consider the vector 

```{r, comment=NA}
set.seed(32020)
st_normal <- rnorm(100)
st_normal
```

Drop the elements corresponds to the positions multiply of 10th (10, 20, 30, ...)

```{r, comment=NA}
st_normal[-seq(1, 100, by=10)]
```