---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 1 

Date: 12 September 2022

Select a `geom_*` function (for example geom_polygon). Make sure that your selected `geom_*` is not explained in any lecture slides. You are allowed to select geoms that I did not explain in https://hellor.netlify.app/slides/l122021.html#46 or new geoms from any other resource.

Fill out the following google sheet with your chosen geom and index number: https://docs.google.com/spreadsheets/d/1OHcEEZdR3LqEfpfMOHRA-VpNCpAbDASSBzR6WICvfT0/edit?usp=sharing

Furthermore, you are not allowed to select a geom that has been already selected by another student in https://docs.google.com/spreadsheets/d/1OHcEEZdR3LqEfpfMOHRA-VpNCpAbDASSBzR6WICvfT0/edit?usp=sharing



Write 

1. a brief explanation of the `geom_*`

2. a ggplot to  demonstrate the application of the selected geom_*

3. R codes and outputs

Output type: You should arrange the content in an RMarkdown file. The output type should be HTML.

Please submit both .Rmd and HTML file.

Due Date: 19 September 2022

## Example

**geom_polygon**

`geom_polygon` can be used to draw boundaries for different regions.

```{r, comment=NA, message=FALSE, warning=FALSE}
library(ggplot2)
library(maps)
library(tidyverse)
world <- map_data("world")
AUS <- world %>% filter(region == "Australia")
ggplot(AUS) + 
   geom_polygon(aes(x = long, y = lat, group = group))
```
