---
title: "STA 326 2.0 Programming and Data Analysis with R"
subtitle: "🎉 "  
author: 
  - ""
  - "Dr Thiyanga Talagala"
date: '`r Sys.Date()`'
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    nature:
      slideNumberFormat: "%current%"
      highlightStyle: github
      highlightLines: true
      ratio: 16:9
      countIncrementalSlides: true
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(
  fig.width=9, fig.height=3.5, fig.retina=3,
  out.width = "100%",
  cache = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE,
  fig.show = TRUE,
  hiline = TRUE
)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
#style_duo_accent(
#  primary_color = "#1381B0",
#  secondary_color = "#FF961C",
#  inverse_header_color = "#FFFFFF"
#)
style_solarized_light(text_font_google   = google_font("Josefin Sans", "400", "400i", "800i", "800"))
#style_mono_light(
#  base_color = "#1c5253",
#  header_font_google = google_font("Josefin Sans"),
#  text_font_google   = google_font("Josefin Sans", "400", "400i", "800i", "800"),
#  code_font_google   = google_font("Fira Mono")
#)
```

## Lecturer-in-charge and Chief Examiner

- Dr Thiyanga S. Talagala

- PhD in Statistics (2019), Monash University, Australia

- BSc (Hons, Special) Statistics, University of Sri Jayewardenepura
      
      - Batch first and Professor R A Dayananda Gold Medalist 
      
- Research interests

      - Data visualization
      
      - Computer vision
      
      - Time series forecasting
      
      - Machine learning and interpretability methods
      
      - Statistical computing
      
- Personal website: https://thiyanga.netlify.app/
---

## Requirements

- Each student is required to join both lectures and exams with a computer with R and RStudio installed.

- Internet connection.



---

---

## Penalties for late submission of assignments

- Each assignment writeup will indicate the deadline for on-time submission. The date and time of the deadline are firm.

- Submitting by the deadline earns the on-time bonus marks. The assignment writeup will specify the bonus if available.

- If you miss the deadline, there may be a grace period for late submissions. The assignment writeup will specify if there is a grace period and how long. You can make a late submission after the deadline and up to the end of the grace period. A penalty is applied for late submissions and the late submissions do not earn the on-time bonus. 

- Late submission penalty

|  Time |  Penalty | On-time bonus  |
|---|:-:|:-:|
| 1 - 2 days  | 0  |  0 |
| 3 - 7 days | -0.5  |  0 |
| 8 - grace period  | -1  |  0 |


- **Submissions are not accepted beyond the grace period.** The grace period is strictly enforced.



---

## Extensions of time for submission

Category A

- An illness that requires the student to miss all of his or her classes for 2 days or more.

Category B

- A period of grief for loss of a family member or friend.

Category C

- A family emergency.

Category D

- Computer breakdown.


**A valid documentation of proof is required prior to the assignment deadline.**
---

## Philosophy driving our late policy has three goals


- Keep things fair.

- Encourage good time-management.

- Encouragement and appreciation of hardworking students.

---


class: center, middle

# Thank you!

Slides created via the R packages:

[**xaringan**](https://github.com/yihui/xaringan)<br>
[gadenbuie/xaringanthemer](https://github.com/gadenbuie/xaringanthemer)

The chakra comes from [remark.js](https://remarkjs.com), [**knitr**](http://yihui.name/knitr), and [R Markdown](https://rmarkdown.rstudio.com).
